Meta-analytic approaches and 'new' effect sizes to account for heterogeneous effects in comparative physiology

### 2021-06-29 MS/ms.Rmd
Just goig through spelling errors

Meta-analysis is a tool that provides comparative physiologists with powerful, unbiased and quantitatively informed answers to topical research questions by synthesising effect sizes across disparate studies. Meta-analysis's core objective and strength is to distill published research results into standardised effect sizes that can be weighted (i.e., by their inverse sampling variance), and which can be compared across broad sets of research designs, study systems and species. Estimating overall effect sizes, and understanding what drives effect variability, provides opportunities to model how organisms will respond to pressing global challenges. Despite this ambition, research designs in comparative physiology can appear, at the outset, as being vastly different to each other (e.g., using different temperatures or treatment dosages). Differences in treatments across studies has led many to believe that meta-analysis is an exercise in comparing "apples with oranges". Here, we dispel this myth by showing how standardised effects sizes can be used in conjunction with powerful multi-level meta-regression models to both account for factors driving differences across studies and make them more comparable. In addition, we derive new effect size measures that provide comparative physiologists with a means to directly make effect measures comparable without the need to resort to more complex statistical models that may be harder to interpret. Our 'new' effect sizes and corresponding sampling variances help physiologists deal with common treatment differeces across studies; allowing researchers to compare both mean differences and associated differences in variance for physiological traits. The new effect sizes we propose, in combination with existing meta-analytic models, will pave the way for comparative physiologists to explore exciting new questions by making results from large-scale data sets collected from the literature more accessible and widely interpretable.

@@

A meta-analysis, being part of systematic review, provides quantitatively informed answers to research questions by using a transparent and reproducible search strategy to synthesise research results across studies and identify sources of variation [@Nakagawa2017; @Gurevitch1999; @Gurevitch2018; @Cooper2009; @Arnqvist1995; @Koricheva2013]. <!--- note meta-analysis itself is not really about transparency etc - it is a stasitical tool - these other bits come in as a part of systeamtic review ---> Meta-analysing independent studies provides greater statistical power and precision than what any individual study on its own would be able to provide on a question - particularly given that most empirical studies are already under-powered in many areas of biology [@Forstmeier2017; @Jennions2003; @Button2013]. By expressing study effects on a common scale (i.e., standardised effect size) a meta-analyst can gain broader insight into the direction and efficacy of a particular treatment effect or the strength of a relationship between two variables of interest [@Gurevitch2018; @Koricheva2013]. Quantitative syntheses, including meta-analyses, have provided comparative physiologists with powerful insights on pressing global challenges - from testing whether physiological plasticity can buffer organisms against climate change [e.g, @Seebacher2015] to the degree to which endocrine disrupting chemicals, such as bisphenol A (BPA), impact aquatic organisms [e.g., @WuSeebacher2020].

Meta-analysis is a well-developed research topic with a long history in statistics, medicine, psychology and ecology [@Gurevitch2018; @Nakagawa2017; @Glass2015]. It has emerged as the 'gold-standard' across various disciplines for quantitative synthesis and is becoming increasingly more prominent in the field of comparative physiology. Formal meta-analyses, as a part of a systematic review, involve a systematic literature-based search, extraction of data from studies meeting specific inclusion criteria, and the subsequent calculation of effects sizes from summary statistics provided by papers. The entire process should follow and report upon this entire process carefully to ensure the meta-analysis within a systematic review, is as transparent and reproducible as possible [i.e., see the Preferred Reporting Items for Systematic Reviews and Meta-Analyses in Ecology and Evolution (PRISMA-EcoEvo) protocol, @ODea2021]. Aside from providing an up-to-date overview of the research question, along with identifying any gaps and biases, there are two major goals of meta-analysis: 1) to provide an overall mean estimate of a treatment effect or relationship and 2) to quantify effect size variance and understand key drivers explaining differences in effects across studies [@Nakagawa2017; @Gurevitch1999; @Gurevitch2018; @Cooper2009; @Koricheva2013]. A critically important feature of meta-analaysis is its ability to account for sampling error [@Koricheva2013]. The inverse of sampling error variance is used to weight studies, accordingly, in the analysis and make overall estimates more precise. While many meta-analyses might employ unweighted analyses, these studies will not gain greater precision on estimates because sampling variance cannot be formally accounted for in the analysis, even if the overall mean itself may be unbiased. <!--- cite Morrissey here - also, the overall mean is unbiased given there is no publication bias or small study effects - this is important as many meta-analyses are likley to suffer from small study effects; my pick sann does not seem like you have talked about publcaiton bias too much - an ablity of meta-analysis to detect publicaiton bias is one of its strength - or probalby the major strength...  --->

Despite the widespread adoption and well established methodological procedures of meta-analysis, it is often criticised for mixing 'apples and oranges' -- in other words, mixing effects from studies that lack comparability [@Gurevitch2018; @Arnqvist1995; @Gallo1979; @Carpenter2020; @Stewart2010; @Glass2015]. Lack of comparability could result from studies differing in experimental design, species, measurement variables and sampling units, which can impact the generalisability of meta-analytic results [@Arnqvist1995; @Stewart2010]. Indeed, study specific differences in comparative physiology are widespread. For example, experimental designs can vary in the temperatures that they apply or in the dosages of chemical modifiers used in experimental treatments [e.g., @WuSeebacher2020]. In addition, they often include a diversity of study species which can vary in responses. To the uninitiated, the concern might not be unfounded given that meta-analyses in ecology and evolution often mix highly heterogeneous effect size estimates, resulting in high variability [@Senior2016]. <!--- this last sentence does not seem to fit well or not completely logical - did you want to say -heterogeniety is normal. Actually Jessica always say in ecology meta-analysis is probalby not that useful; meta-regression is what we are most interested - I think yo cna make a similar point or you are already making one but make it clearer  - the first sentence is a good one but rather than focusing on "contolling" for hetero but finding out what causes hetero should our (comparative phsiologists') main aim for meta-analysis. I know you are probably intending to say something like this but not coming through at the moment. Also, I note that "review generated evidence" only comes from meta-regression not meta-analysis = intercept models - as study general evidnece can be equivalent to meta-analytic results --->

The goal of our review is to briefly overview meta-analytic practices and reassure comparative physiologists that heterogeneity among effect sizes can be comfortably dealt with by modifying existing effect sizes measures and using multilevel meta-regression models. We first conduct a survey of published meta-analyses in the field of comparative physiology to gauge methodological and meta-analytic approaches - assessing what types of effect sizes are commonly used, how often effect size estimates are impacted by continuous treatment differences and whether meta-analytic models (i.e., models that account for sampling variance) are commonly applied. We then introduce new effect sizes and their associated sampling variance to provide comparative physiologists opportunities to explicitly incorporate treatment differences into the effect size estimates to ease their interpretation. <!--- see my comments above - they are not really new effec sizes - just formalising we can chat about this lnRR, lnVR and lnCVR and Q10 all exisit already ---> We then describe how more complex treatment differences, such as dosage differences, which may have non-linear effects, can be accommodated using multilevel meta-regression models. We hope to show how expanding the meta-analytic toolkit for comparative physiologists will provide new opportunities to address global scale problems that attempt to understand how organisms will cope with rapidly changing environments in the future.

@@

<!---  Dan syas: I'm thinking now, in re-reading this, that we could use a slightly expanded version of paragraph 4 taht describes this problem in a little more detail. This is likely the place for a figure, which I am having some trouble envisaging where one is best placed. It just otherwise seems like a bit of a jump to otherwise head straight into the survey. Here we should also talk about quality control, the need to be transparent and have careful reporting. This is a big issue in medicine. Talked about a lot. Integrate blurb below…try to take a balanced approach. I will add an introduction bit to this paragraph below to put into context…. xxxx --->

In the words of Gene Glass, the father of meta-analysis: "Of course it mixes apples and oranges; in the study of fruit nothing else is sensible; comparing apples and oranges is the only endeavor worthy of true scientists; comparing apples to apples is trivial" [p.g., 224, @Glass2015]. @Glass2015's intention was to simply highlight that whether comparing apples and oranges is a good idea simply boils down to the specific question of interest and the population one wishes to make conclusions about. If we are interested in gaining insight into 'fruit' then synthesising apples and oranges is the only thing that makes sense. <!---  as said, comparative analysis by definition, compares applies and oranges - = different species - I guess you are mainly talking about mixing different methods ones - this is where I said earlier - focusing on explaining hetero comes in. Also, mixing different quality of studies are often referred to as "Garbage in and garbage out" ---> Instead, if we are interested in generalising only to apples then mixing fruit will not be ideal. We tend to agree with @Glass2015, and his sentiment is of particular relevance to comparative physiology. For example, if we want to understand the impact of a pollutant on reproduction in aquatic organisms then including aquatic mammals and fish would provide answers to this general question, even if their reproductive biology and physiology are known to differ. By synthesizing diverse study outcomes, it allows us to not just report on overall effects (if the question is relevant to ask), but understand critical features of the literature, study systems (e.g., different effects of pollutants on fish and mammal reproduction) and approaches that can explain the diversity of effect outcomes observed. In doing so, it provides a richer set of conclusions that can draw attention to important sources of variability that might have otherwise been overlooked. Having said that, it is still very important to understand the limitations and diversity of studies included within a meta-analysis - an important reason why transparency and reproducibility are such a prominent feature of meta-analysis (or ideally should be) [@ODea2021]. The validity of general conclusions to certain questions (e.g., what is the overall effect of a pollutant on reproduction?) will inevitably depend on the homogeneity of effects being synthesized. <!--- this is know as putting one number on so many different kinds of studies - but of course, this may not be that interesting apart from a starting point for subsequent meta-regressoin analyses ---> For example, given the very different physiology of aquatic mammals and fish, understanding an overall combined effect on reproduction may not be particularly relevant, and indeed may be misleading. It is important to consider these questions when conducting meta-analyses.

Of course, there are certain types of heterogeneity that transcend different questions. In other words, if we compare different types of 'apples' is it still fair to synthesize these studies to understand overall effects? This situation might be analogous to a meta-analysis in comparative physiology that attempts to understand the effects of temperature on fish growth. <!--- taxnomically speaking - comparing different applies are comparing different populations of the same spieces of fish so not quite sure whether this is quite correct - I am being of course a bit pedentic---> Studies might rear populations of fish under different temperatures, but in the end, they are still all fish, governed by the same thermodynamic effects on growth. In this situation, we might still be interested in understanding overall effects, but we may question whether such a conclusion is sensible given the diversity of temperatures applied. Such a study does not preclude us from also exploring important moderators or predictors of variability in effects across studies.....

@@

As discussed above, it is clear that different types of questions and experimental approaches will be more or less susceptable to the 'apples and oranges' problem. <!--- not sure whehter it is a good idea to hightlight it as a problem as said earlier - explaining hetero is what's interesting and even the main porupose of meta-analysis for comparative physiologiests - they are not the problem but they are what we are going to invesitgate. I do not think 'suspetible' is a good word here given the average hetero = > 90% - you are nearly 100% guranteed to have hetero and this is what's interesting ---> As such, before embarking on our path to 'meta-analytic enlightenment', we conducted a literature survey to understand how widespread the problem may be in comparative physiology and gauge the ways in which it is currently being addressed, if at all. To accomplish this goal, we searched for meta-analyses and comparative analyses that tackled questions using physiological data compiled from animal studies across multiple species. We were specifically interested in asking: 1) what type of effect size was used within the quantitative synthesis? 2) based on the methods provided, was the effect size likely susceptible to variation in how treatments are applied across studies? 3) If so, was the variation resulting from treatment differences accounted for in the effect size or during analysis?, and finally, 4) how were effect sizes analysed in the quantitative synthesis, were these weighted analyses, and what program was used for analysis? We also recorded whether publication bias and sensitivity analyses were conducted, if data were publicly available, where it was stored, and if so, if it was usable (i.e., in a machine readable format).

@@

All searches were conducted in Scopus between May 16 - 18, 2021. Briefly, we first did a series of pilot searches (detailed in Supplementary Materials) to refine our search query. We searched for studies that included: 'meta-analysis', 'meta regression' , 'comparative analysis', 'comprehensive analysis' 'global analysis' or 'macrophysiology' in their title and abstract, restricting our search to key comparative physiology journals (deatiled in Table X - Supplementary Materials) within the last 6 years. We expected this to provide a contemporay overview of quantitative syntheses in comparative physiology while making the survey and screening feasible. We awknowledge that this may have missed some important, and influential, quantitative synthese in the field of comparative physiology, but likely provided a random sample to inform on the questions we sought to address (see below).

We included quantitative syntheses, including meta-analyses (*sensu stricto*), that were focused on comparative physiologcial questions only. Narrative reviews and comments were not considered because they did not collate quantitative data from the literature. Studies needed to have collected physiological trait data from the literature including traits, such as hormones, thermal physiology (CTmax, CTmin, thermal preferences, thermal breadth, LT50, lethal thresholds), metabolic rates (routine, maximum, and minimum metabolic rate, aerobic scope), enzyme reaction rates and concentrations, oxidative stress and consequences (e.g., DNA damage, teleomere attrition), performance measurements (e.g., heart rates, sprint speed, swim speed, growth rates), body composition (e.g., fatty acid profiles) and immune traits (e.g., cell function, antibodies, ect.). Studies were also included if they measured the relationship between physiological traits and body size (e.g., metabolic scaling), reproductive allocation, or survival / mortality across taxa. Studies on behavior were only included if the synthesis was explicitly interested in the proximate physiological drivers of behavioral variation. We also included meta-analyses that quantified plasticity in physiological traits across taxa for the purpose of understanding adaptive plasticity in response to environmental variation. While many studies in comparative physiology collect data as part of a comparative analysis, we focus here only on studies collating data from existing literature because this is the major data source for quantitative synthesis. Quantitative syntheses interested in phenology, biomechanics, sexual size dimorphism (i.e., SSD, except if the study was interested in sexual differences in physiological traits), those focused on validating physiological methodology, or characterising population genetic structure, gene transcription and quantitative genetics were excluded. There were also meta-analyses that focused on colour evolution and contrast which varied in their goals but were most often not focused on physiological questions. As such, these studies were also exclude. We restricted our focus to animal studies because plant synthesis papers were often only focused on ecological (e.g., quantified changes in abundance and distribution of species) or agricultural questions (e.g., studies on soil microbial function, nitrogen and phosphorus capture, methane output and greenhouse gas emissions).

@@

Meta-analysis requires comparative physiologists to make use of effect sizes that are suitable for answering their specific question. The key features of all effect sizes used in meta-analyses are that they are: 1) statistical parameters that have been placed on a common scale so that they can be compared across studies and 2) that they have some associated measure of precision (i.e., sampling variance)[@Koricheva2013; @Borenstein2009; @Schmid2021]. Effect sizes in comparative physiology are quite variable; they can be the mean difference between two groups (e.g., difference in mean between a control and treatment group - see @WuSeebacher2020), the slope of a regression (e.g., between body size and metabolic rate), a correlation between two variables, or even the raw mean (or variance) itself (e.g., comparing CTmax or CTmin across species - see @Sunday2014). In many cases, multiple types of effect sizes might be used depending on the question. For example, studies interested in the effect of temperature on some physiological trait that have measured the trait across studies experimentally manipulating temperature could analyse the mean 'effect' by creating contrast-based effect sizes or by analysing the mean in each temperature treatment. The choice as to which one to use will boil down to the types of data available from primary studies, the specific question of interest and whether studies are observational or experimental in nature. Considering our temperature example above, if we are interested in the magnitude of change between temperature treatments, we may want to use a contrast-based effect size, such as the log response ratio $\ln{RR}$. If we are instead interested in how the mean of some physiological trait changes with tempearture we may instead model the mean itself. In the latter scenerio, so long as all the means are measured in the same units (e.g., ; or can be converted to the same units) then the mean may provide the desired answer to a given question. We provide details on different effect measures and associated sample sizes in Table \@ref(tab:tablea).

@@

A very common occurrence in comparative physiology is to manipulate the temperature organisms experience and measure some physiological rate (e.g., metabolic rate). Using normal effect size measures (see above) would result in the effect sizes varying in the types of temperatures applied within and across across studies contributing to 'comparability' issues. As such, it is often of interest to standardsise effects by temperature. One common effect size measure already used in comparative physiology to compare physiological rates is $Q_{10}$ [e.g., @Rodgers2021; @Seebacher2015; @Havird2020], which describes the multiplicative change in physiological rates across a 10°C temperature change. Higher $Q_{10}$ values indicate larger changes in physiological rates. Currently, however, we do not have associated sampling variances derived for $Q_{10}$ making it challenging to make use of the full power of weighted meta-analytic models. Interestingly, $Q_{10}$ can be seen as a variant of *lnRR*, meaning that we can derive a $Q_{10}$ based effect size and sampling error using the well known mathematical properties of *lnRR*. This opens up the meta-analytic toolbox and improves upon our ability to account for well known sources of non-independence that are typical in meta-analysis [@Lajeunesse2011; @Noble2017].

@@

<!--- As said before you want to play up the aspect of investigaing heterogeneity rather than controlling for it - controlling for it seems to mean you still want to know the overall effect after getting rid of every confunders. Whilst it is also waht we want, we really want to explain the hetero --->

@@

Heterogeneity among effects that result from known differences in how treatments are applied can also be dealt with using multilevel meta-regression models. Multilevel meta-regression models include all the same random effects that we have already discussed, but they also include fixed effects (i.e., predictors or moderators) that attempt to explain changes in the mean effect size. In otherwords, we can modify our multilevel meta-analytic model to turn it into a multilevel meta-regression model as follows:

>>>>

Meta-analysis is a tool that provides comparative physiologists with powerful, unbiased and quantitatively informed answers to topical research questions by synthesising effect sizes across disparate studies. Meta-analysis's core objective and strength is to distill published research results into standardised effect sizes that can be weighted (i.e., by their inverse sampling variance), and which can be compared across broad sets of research designs, study systems and species. Estimating overall effect sizes, and understanding what drives effect variability, provides opportunities to model how organisms will respond to pressing global challenges. Despite this ambition, research designs in comparative physiology can appear, at the outset, as being vastly different to each other (e.g., using different temperatures or treatment dosages). Differences in treatments across studies has led many to believe that meta-analysis is an exercise in comparing "apples with oranges". Here, we dispel this myth by showing how standardised effects sizes can be used in conjunction with powerful multi-level meta-regression models to both account for factors driving differences across studies and make them more comparable. In addition, we derive new effect size measures that provide comparative physiologists with a means to directly make effect measures comparable without the need to resort to more complex statistical models that may be harder to interpret. Our 'new' effect sizes and corresponding sampling variances help physiologists deal with common treatment differences across studies; allowing researchers to compare both mean differences and associated differences in variance for physiological traits. The new effect sizes we propose, in combination with existing meta-analytic models, will pave the way for comparative physiologists to explore exciting new questions by making results from large-scale data sets collected from the literature more accessible and widely interpretable.

@@

A meta-analysis, being part of systematic review, provides quantitatively informed answers to research questions by using a transparent and reproducible search strategy to synthesise research results across studies and identify sources of variation [@Nakagawa2017; @Gurevitch1999; @Gurevitch2018; @Cooper2009; @Arnqvist1995; @Koricheva2013]. <!--- note meta-analysis itself is not really about transparency etc - it is a stasitical tool - these other bits come in as a part of systematic review ---> Meta-analysing independent studies provides greater statistical power and precision than what any individual study on its own would be able to provide on a question - particularly given that most empirical studies are already under-powered in many areas of biology [@Forstmeier2017; @Jennions2003; @Button2013]. By expressing study effects on a common scale (i.e., standardised effect size) a meta-analyst can gain broader insight into the direction and efficacy of a particular treatment effect or the strength of a relationship between two variables of interest [@Gurevitch2018; @Koricheva2013]. Quantitative syntheses, including meta-analyses, have provided comparative physiologists with powerful insights on pressing global challenges - from testing whether physiological plasticity can buffer organisms against climate change [e.g, @Seebacher2015] to the degree to which endocrine disrupting chemicals, such as bis phenol A (BPA), impact aquatic organisms [e.g., @WuSeebacher2020].

Meta-analysis is a well-developed research topic with a long history in statistics, medicine, psychology and ecology [@Gurevitch2018; @Nakagawa2017; @Glass2015]. It has emerged as the 'gold-standard' across various disciplines for quantitative synthesis and is becoming increasingly more prominent in the field of comparative physiology. Formal meta-analyses, as a part of a systematic review, involve a systematic literature-based search, extraction of data from studies meeting specific inclusion criteria, and the subsequent calculation of effects sizes from summary statistics provided by papers. The entire process should follow and report upon this entire process carefully to ensure the meta-analysis within a systematic review, is as transparent and reproducible as possible [i.e., see the Preferred Reporting Items for Systematic Reviews and Meta-Analyses in Ecology and Evolution (PRISMA-EcoEvo) protocol, @ODea2021]. Aside from providing an up-to-date overview of the research question, along with identifying any gaps and biases, there are two major goals of meta-analysis: 1) to provide an overall mean estimate of a treatment effect or relationship and 2) to quantify effect size variance and understand key drivers explaining differences in effects across studies [@Nakagawa2017; @Gurevitch1999; @Gurevitch2018; @Cooper2009; @Koricheva2013]. A critically important feature of meta-analysis is its ability to account for sampling error [@Koricheva2013]. The inverse of sampling error variance is used to weight studies, accordingly, in the analysis and make overall estimates more precise. While many meta-analyses might employ unweighted analyses, these studies will not gain greater precision on estimates because sampling variance cannot be formally accounted for in the analysis, even if the overall mean itself may be unbiased. <!--- cite Morrissey here - also, the overall mean is unbiased given there is no publication bias or small study effects - this is important as many meta-analyses are likley to suffer from small study effects; my pick sann does not seem like you have talked about publcaiton bias too much - an ability of meta-analysis to detect publication bias is one of its strength - or probably the major strength...  --->

Despite the widespread adoption and well established methodological procedures of meta-analysis, it is often criticised for mixing 'apples and oranges' -- in other words, mixing effects from studies that lack comparability [@Gurevitch2018; @Arnqvist1995; @Gallo1979; @Carpenter2020; @Stewart2010; @Glass2015]. Lack of comparability could result from studies differing in experimental design, species, measurement variables and sampling units, which can impact the generalisability of meta-analytic results [@Arnqvist1995; @Stewart2010]. Indeed, study specific differences in comparative physiology are widespread. For example, experimental designs can vary in the temperatures that they apply or in the dosages of chemical modifiers used in experimental treatments [e.g., @WuSeebacher2020]. In addition, they often include a diversity of study species which can vary in responses. To the uninitiated, the concern might not be unfounded given that meta-analyses in ecology and evolution often mix highly heterogeneous effect size estimates, resulting in high variability [@Senior2016]. <!--- this last sentence does not seem to fit well or not completely logical - did you want to say -heterogeneity is normal. Actually Jessica always say in ecology meta-analysis is probably not that useful; meta-regression is what we are most interested - I think yo can make a similar point or you are already making one but make it clearer  - the first sentence is a good one but rather than focusing on "controlling" for hetero but finding out what causes hetero should our (comparative physiologists') main aim for meta-analysis. I know you are probably intending to say something like this but not coming through at the moment. Also, I note that "review generated evidence" only comes from meta-regression not meta-analysis = intercept models - as study general evidence can be equivalent to meta-analytic results --->

The goal of our review is to briefly overview meta-analytic practices and reassure comparative physiologists that heterogeneity among effect sizes can be comfortably dealt with by modifying existing effect sizes measures and using multilevel meta-regression models. We first conduct a survey of published meta-analyses in the field of comparative physiology to gauge methodological and meta-analytic approaches - assessing what types of effect sizes are commonly used, how often effect size estimates are impacted by continuous treatment differences and whether meta-analytic models (i.e., models that account for sampling variance) are commonly applied. We then introduce new effect sizes and their associated sampling variance to provide comparative physiologists opportunities to explicitly incorporate treatment differences into the effect size estimates to ease their interpretation. <!--- see my comments above - they are not really new effect sizes - just formalising we can chat about this lnRR, lnVR and lnCVR and Q10 all exisit already ---> We then describe how more complex treatment differences, such as dosage differences, which may have non-linear effects, can be accommodated using multilevel meta-regression models. We hope to show how expanding the meta-analytic toolkit for comparative physiologists will provide new opportunities to address global scale problems that attempt to understand how organisms will cope with rapidly changing environments in the future.

@@

<!---  Dan says: I'm thinking now, in re-reading this, that we could use a slightly expanded version of paragraph 4 that describes this problem in a little more detail. This is likely the place for a figure, which I am having some trouble envisaging where one is best placed. It just otherwise seems like a bit of a jump to otherwise head straight into the survey. Here we should also talk about quality control, the need to be transparent and have careful reporting. This is a big issue in medicine. Talked about a lot. Integrate blurb below…try to take a balanced approach. I will add an introduction bit to this paragraph below to put into context…. xxxx --->

In the words of Gene Glass, the father of meta-analysis: "Of course it mixes apples and oranges; in the study of fruit nothing else is sensible; comparing apples and oranges is the only endeavor worthy of true scientists; comparing apples to apples is trivial" [p.g., 224, @Glass2015]. @Glass2015's intention was to simply highlight that whether comparing apples and oranges is a good idea simply boils down to the specific question of interest and the population one wishes to make conclusions about. If we are interested in gaining insight into 'fruit' then synthesising apples and oranges is the only thing that makes sense. <!---  as said, comparative analysis by definition, compares applies and oranges - = different species - I guess you are mainly talking about mixing different methods ones - this is where I said earlier - focusing on explaining hetero comes in. Also, mixing different quality of studies are often referred to as "Garbage in and garbage out" ---> Instead, if we are interested in generalising only to apples then mixing fruit will not be ideal. We tend to agree with @Glass2015, and his sentiment is of particular relevance to comparative physiology. For example, if we want to understand the impact of a pollutant on reproduction in aquatic organisms then including aquatic mammals and fish would provide answers to this general question, even if their reproductive biology and physiology are known to differ. By synthesizing diverse study outcomes, it allows us to not just report on overall effects (if the question is relevant to ask), but understand critical features of the literature, study systems (e.g., different effects of pollutants on fish and mammal reproduction) and approaches that can explain the diversity of effect outcomes observed. In doing so, it provides a richer set of conclusions that can draw attention to important sources of variability that might have otherwise been overlooked. Having said that, it is still very important to understand the limitations and diversity of studies included within a meta-analysis - an important reason why transparency and reproducibility are such a prominent feature of meta-analysis (or ideally should be) [@ODea2021]. The validity of general conclusions to certain questions (e.g., what is the overall effect of a pollutant on reproduction?) will inevitably depend on the homogeneity of effects being synthesized. <!--- this is know as putting one number on so many different kinds of studies - but of course, this may not be that interesting apart from a starting point for subsequent meta-regression analyses ---> For example, given the very different physiology of aquatic mammals and fish, understanding an overall combined effect on reproduction may not be particularly relevant, and indeed may be misleading. It is important to consider these questions when conducting meta-analyses.

Of course, there are certain types of heterogeneity that transcend different questions. In other words, if we compare different types of 'apples' is it still fair to synthesize these studies to understand overall effects? This situation might be analogous to a meta-analysis in comparative physiology that attempts to understand the effects of temperature on fish growth. <!--- taxnomically speaking - comparing different applies are comparing different populations of the same species of fish so not quite sure whether this is quite correct - I am being of course a bit pedantic---> Studies might rear populations of fish under different temperatures, but in the end, they are still all fish, governed by the same thermodynamic effects on growth. In this situation, we might still be interested in understanding overall effects, but we may question whether such a conclusion is sensible given the diversity of temperatures applied. Such a study does not preclude us from also exploring important moderators or predictors of variability in effects across studies.....

@@

As discussed above, it is clear that different types of questions and experimental approaches will be more or less susceptible to the 'apples and oranges' problem. <!--- not sure whether it is a good idea to highlight it as a problem as said earlier - explaining hetero is what's interesting and even the main purpose of meta-analysis for comparative physiologists - they are not the problem but they are what we are going to investigate. I do not think 'susceptible' is a good word here given the average hetero = > 90% - you are nearly 100% guaranteed to have hetero and this is what's interesting ---> As such, before embarking on our path to 'meta-analytic enlightenment', we conducted a literature survey to understand how widespread the problem may be in comparative physiology and gauge the ways in which it is currently being addressed, if at all. To accomplish this goal, we searched for meta-analyses and comparative analyses that tackled questions using physiological data compiled from animal studies across multiple species. We were specifically interested in asking: 1) what type of effect size was used within the quantitative synthesis? 2) based on the methods provided, was the effect size likely susceptible to variation in how treatments are applied across studies? 3) If so, was the variation resulting from treatment differences accounted for in the effect size or during analysis?, and finally, 4) how were effect sizes analysed in the quantitative synthesis, were these weighted analyses, and what program was used for analysis? We also recorded whether publication bias and sensitivity analyses were conducted, if data were publicly available, where it was stored, and if so, if it was usable (i.e., in a machine readable format).

@@

All searches were conducted in Scopus between May 16 - 18, 2021. Briefly, we first did a series of pilot searches (detailed in Supplementary Materials) to refine our search query. We searched for studies that included: 'meta-analysis', 'meta regression' , 'comparative analysis', 'comprehensive analysis' 'global analysis' or 'macro physiology' in their title and abstract, restricting our search to key comparative physiology journals (detailed in Table X - Supplementary Materials) within the last 6 years. We expected this to provide a contemporary overview of quantitative syntheses in comparative physiology while making the survey and screening feasible. We acknowledge that this may have missed some important, and influential, quantitative synthesis in the field of comparative physiology, but likely provided a random sample to inform on the questions we sought to address (see below).

We included quantitative syntheses, including meta-analyses (*sensu stricto*), that were focused on comparative physiological questions only. Narrative reviews and comments were not considered because they did not collate quantitative data from the literature. Studies needed to have collected physiological trait data from the literature including traits, such as hormones, thermal physiology (CTmax, CTmin, thermal preferences, thermal breadth, LT50, lethal thresholds), metabolic rates (routine, maximum, and minimum metabolic rate, aerobic scope), enzyme reaction rates and concentrations, oxidative stress and consequences (e.g., DNA damage, teleomere attrition), performance measurements (e.g., heart rates, sprint speed, swim speed, growth rates), body composition (e.g., fatty acid profiles) and immune traits (e.g., cell function, antibodies, ect.). Studies were also included if they measured the relationship between physiological traits and body size (e.g., metabolic scaling), reproductive allocation, or survival / mortality across taxa. Studies on behavior were only included if the synthesis was explicitly interested in the proximate physiological drivers of behavioral variation. We also included meta-analyses that quantified plasticity in physiological traits across taxa for the purpose of understanding adaptive plasticity in response to environmental variation. While many studies in comparative physiology collect data as part of a comparative analysis, we focus here only on studies collating data from existing literature because this is the major data source for quantitative synthesis. Quantitative syntheses interested in phenology, bio-mechanics, sexual size dimorphism (i.e., SSD, except if the study was interested in sexual differences in physiological traits), those focused on validating physiological methodology, or characterising population genetic structure, gene transcription and quantitative genetics were excluded. There were also meta-analyses that focused on colour evolution and contrast which varied in their goals but were most often not focused on physiological questions. As such, these studies were also exclude. We restricted our focus to animal studies because plant synthesis papers were often only focused on ecological (e.g., quantified changes in abundance and distribution of species) or agricultural questions (e.g., studies on soil microbial function, nitrogen and phosphorus capture, methane output and greenhouse gas emissions).

@@

Meta-analysis requires comparative physiologists to make use of effect sizes that are suitable for answering their specific question. The key features of all effect sizes used in meta-analyses are that they are: 1) statistical parameters that have been placed on a common scale so that they can be compared across studies and 2) that they have some associated measure of precision (i.e., sampling variance)[@Koricheva2013; @Borenstein2009; @Schmid2021]. Effect sizes in comparative physiology are quite variable; they can be the mean difference between two groups (e.g., difference in mean between a control and treatment group - see @WuSeebacher2020), the slope of a regression (e.g., between body size and metabolic rate), a correlation between two variables, or even the raw mean (or variance) itself (e.g., comparing CTmax or CTmin across species - see @Sunday2014). In many cases, multiple types of effect sizes might be used depending on the question. For example, studies interested in the effect of temperature on some physiological trait that have measured the trait across studies experimentally manipulating temperature could analyse the mean 'effect' by creating contrast-based effect sizes or by analysing the mean in each temperature treatment. The choice as to which one to use will boil down to the types of data available from primary studies, the specific question of interest and whether studies are observational or experimental in nature. Considering our temperature example above, if we are interested in the magnitude of change between temperature treatments, we may want to use a contrast-based effect size, such as the log response ratio $\ln{RR}$. If we are instead interested in how the mean of some physiological trait changes with temperature we may instead model the mean itself. In the latter scenario, so long as all the means are measured in the same units (e.g., ; or can be converted to the same units) then the mean may provide the desired answer to a given question. We provide details on different effect measures and associated sample sizes in Table \@ref(tab:tablea).

@@

A very common occurrence in comparative physiology is to manipulate the temperature organisms experience and measure some physiological rate (e.g., metabolic rate). Using normal effect size measures (see above) would result in the effect sizes varying in the types of temperatures applied within and across across studies contributing to 'comparability' issues. As such, it is often of interest to standardize effects by temperature. One common effect size measure already used in comparative physiology to compare physiological rates is $Q_{10}$ [e.g., @Rodgers2021; @Seebacher2015; @Havird2020], which describes the multiplicative change in physiological rates across a 10°C temperature change. Higher $Q_{10}$ values indicate larger changes in physiological rates. Currently, however, we do not have associated sampling variances derived for $Q_{10}$ making it challenging to make use of the full power of weighted meta-analytic models. Interestingly, $Q_{10}$ can be seen as a variant of *lnRR*, meaning that we can derive a $Q_{10}$ based effect size and sampling error using the well known mathematical properties of *lnRR*. This opens up the meta-analytic toolbox and improves upon our ability to account for well known sources of non-independence that are typical in meta-analysis [@Lajeunesse2011; @Noble2017].

@@

<!--- As said before you want to play up the aspect of investigaing heterogeneity rather than controlling for it - controlling for it seems to mean you still want to know the overall effect after getting rid of every confunders. Whilst it is also what we want, we really want to explain the hetero --->

@@

Heterogeneity among effects that result from known differences in how treatments are applied can also be dealt with using multilevel meta-regression models. Multilevel meta-regression models include all the same random effects that we have already discussed, but they also include fixed effects (i.e., predictors or moderators) that attempt to explain changes in the mean effect size. In other words, we can modify our multilevel meta-analytic model to turn it into a multilevel meta-regression model as follows:
